### 2023-09-25 paper/draft.tex
Initial Overleaf Import

%\title{Motivating selective classification for document classification} %\project
%\maketitle
\begin{center}{\noindent\Large\bf Defining Selective Multi-Class Classification}\end{center}

\section{AURC definition}

Given a multi-class classifier $f: \mathcal{X} \rightarrow \Delta^{k} \times [0,1]$ for a $k$ class problem with a reject option, where $\Delta$ is a probability simplex. %assumes additional logit; theta form part

\noindent Origin: \url{https://arxiv.org/pdf/1805.08206.pdf}

\begin{align}
    \operatorname{AURC}(f) := \mathbb{E}_{x \sim p_x} \left[  \frac{\mathbb{E}_{(\tilde{x},\tilde{y})\sim p_{xy}}[\ell([f(\tilde{x})]_{1:k},\tilde{y}) \mathbb{I}[f(\tilde{x})_{k+1} > f(x)_{k+1}]]}{\mathbb{E}_{\tilde{x}\sim p_{x}}[ \mathbb{I}[f(\tilde{x})_{k+1} > f(x)_{k+1}]]} \right] \\
    \approx \frac{1}{n} \sum_{h=1}^n \frac{\frac{1}{n-1} \sum_{i\neq h} \ell([f(x_i)]_{1:k},y_i) \mathbb{I}[f(x_i)_{k+1} > f(x_h)_{k+1}]}{\frac{1}{n-1} \sum_{i\neq h} \mathbb{I}[f(x_i)_{k+1} > f(x_h)_{k+1}]}
\end{align}

\section{Example \AURC}

\newpage

\begin{center}{\noindent\Large\bf Motivating Selective Document Classification}\end{center}

\begin{uuredbox}
This version (18/09) has been stated in terms of the context and motivation, not yet as properly formulated definitions; bear with me on this :) 
\end{uuredbox}

\section{Context}

Document classification (\DC) is a quintessential task in intelligent document processing workflows. In practice, it is characterized as a multi-class classification problem with typically high cardinality (100-500 mutually exclusive document types) \cite{vanlandeghem2023document}. Production-level \DC{} models operate in a high-accuracy regime (>80\%) and are expected to be both \textit{classification-calibrated} \cite[Definition 1]{ni2019calibration} (top-1 calibration \cite{naeini2015obtaining,guo2017calibration,vaicenavicius2019evaluating}) and \textit{rejection-calibrated} \cite[Definition 2]{ni2019calibration}. As the predicted document type dictates further processing of key information extraction (KIE), typically specified according to static formats (e.g., a receipt for expenditure specifies different information than a car collision form), this primary step in the pipeline is crucial.

%not SOTA though

\section{Motivation}

There is a lot of interest in reducing overall risk while maintaining the highest possible target coverage in large multi-class classification problems. Naturally, rejecting low-confidence, uncertain samples for human intervention is preferred to trusting a classifier. 
This is the target of interest, whereas we train our models with a differentiable objective (commonly cross-entropy loss $\ell_{\mathrm{CE}}$) that might not be aligned. \cite{geifman2017selective,geifman2018bias} point out that there is no consensus on the appropriate performance measure for scoring functions.
In practice, the classes can be imbalanced and might have a higher associated misclassification cost, although users are not likely to quantify class-specific costs for a label set of high cardinality. %there is also a link with fairness, as some document types might be more privy to processing 
An additional practical problem exists in finding an appropriate set of class-specific thresholds that overall minimizes the target risk and maximizes a high (>70\%) target coverage.

In the \DC{} task setting, it is more sensible to target low-risk levels, instead of targeting midpoint coverage rates. Therefore, I propose to research a surrogate loss function with additional control over the risk, incentivizing the model to prioritize the true rejection rate.

%TODO: formulate correctly as definitions

\section{Background}

\subsection{Calibration}

\todo[inline]{Further establish link/difference to calibration}

\noindent \begin{uuyellowbox}[title=Top-1 calibration does not imply good failure prediction \cite{zhu2022rethinking}]

    Consider two probabilistic classifiers, each providing max softmax probability (MSP) as the confidence scoring function (CSF).\newline

    Classifier A: 90\% accuracy - varying MSP ranging from 0.8-1

    Classifier B: 92\% accuracy - MSP always 0.92\newline

    Which classifier would you choose in practice (keeping in mind the goals of selective classification)?

\end{uuyellowbox}

\subsection{Selective classification}

It has been referred to as confidence ranking in the context of \textit{iid} misprediction classification or failure/error prediction \cite{jaeger2023a}. For a good overview, check the survey \cite{zhang2023survey}.

\noindent Popular \textbf{evaluation metrics} include:
\begin{enumerate}
    \item \AUROC{} (Area Under the ROC Curve)
    \item \AURC{} (Area Under the Risk Coverage Curve) \citep{geifman2017selective,jaeger2023a} 
    \item \EAURC{} (discounting accuracy and normalization) \citep{geifman2018bias}
\end{enumerate}

\noindent Similar to calibration research, there are \textbf{I. train-time} methods involving surrogate losses, architecture \& optimization extensions, and \textbf{II. training-free} methods.
\todo[inline]{Add related work, \href{https://openreview.net/pdf?id=K1DdnjL6p7}{recent reference}}

\noindent Important \textbf{theoretical result}:

Impossibility of "separation-based" calibrated surrogates for multi-class classification \cite[Theorem 4]{ni2019calibration}, making it a must-read. If I understand it well enough, it posits that $\ell_{\mathrm{CE}}$ is the best available choice.

Another important related work \href{https://arxiv.org/pdf/2208.12084.pdf}{Calibrated Selective Classification} \cite{fisch2022calibrated} proposes a trainable objective for selective calibration based on MMCE \cite{kumar2018trainable} \footnote{There is \href{https://github.com/ajfisch/calibrated-selective-classification}{code}}.

\subsection{Knowledge distillation}

Teacher-student training involves soft labels, which AFAIK has not been investigated together with selective classification. Matthew suggested swapping out $\ell_{\mathrm{CE}}$ for a proposed surrogate (\Cref{sec:method}) while fine-tuning the teacher and/or student network. 
Recent work \cite{galil2023what} showed that models trained with knowledge distillation demonstrated improved \AURC{}.

\section{Method}\label{sec:method}

\subsubsection{Prior methods}

\cite{geifman2017selective,geifman2018bias} provides the seminal paper with a computationally heavy method (ad-hoc loss and architecture extension) to control coverage and risk at training time.
%for training a neural network with three heads: a classification head, and auxiliary head, and one head to estimate the confidence of the classification decision.

Regarding I. it is hard to establish a clear winning method (in terms of \AURC) over MSP, the literature has shown that differences are often not reproducible\footnote{\href{https://github.com/LayneH/SAT-selective-cls/issues/3\#issuecomment-1640151933}{Discussion} ongoing.} or not statistically significant \cite{jaeger2023a}.
The current (II.) state-of-the-art \cite{lin2022scrib} treats it as a set-based prediction with a post-hoc multi-thresholding algorithm that reduces overall risk. They devised an algorithm based on coordinate ascent (check 1 variable, keep others fixed) with some efficiency optimization at the lower level of class-specific thresholding.

\todo[inline]{Describe the optimal solution with time/space complexity (Big-Oh) and naive implementation}

\todo[inline]{Describe a baseline solution}

% \subsection{Task description}

% \subsection{Method contributions}

% \begin{itemize}
% \item 
% \end{itemize}

\section{Experiments}

I have created a benchmark (compatible with HuggingFace) for \KD{} on \DC{}.

\subsection{Selective Training Methods}

For teacher fine-tuning, the following (I.) methods have been implemented:

\begin{enumerate}
    \item $\ell_{\mathrm{CE}}$ baseline 
    \item Confidence Ranking Loss \cite{moon2020confidence}
    \item LogitNorm Loss \cite{wei2022mitigating}
    \item Deep Gamblers Loss \cite{liu2019deep}
    \item Symmetric Cross-Entropy \cite{wang2019symmetric}
    \item Entropy Regularization \cite{feng2022stop} 
    \item Self-adaptive Training \cite{huang2020self}
    \item Intermediate Checkpointing \cite{wang2022efficient}
\end{enumerate}

Other interesting (II.) training-free methods to be investigated (\#TODO): 
\begin{itemize}
    \item DOCTOR \cite{granese2021doctor}
    \item Gini \cite{gomes2023a}
    \item SCRIBR \cite{lin2022scrib}
\end{itemize}

\subsection{Distillation Methods}

Current \KD{} methods are supported: 
\begin{itemize}
\item  Response/logits-based
\begin{enumerate}
    \item Cross-entropy with KD hyperparameters ($\alpha$ and $\tau$) \cite{hinton2015distilling}
    \item Mean Squared Error \cite{kim2021comparing}
    \item NKD \cite{yang2023knowledge}
    %\item Need something different for FasterRCNN
    %\item (Hyperparameter grid search run for all)
\end{enumerate}
\item  Features-based
\begin{enumerate}
    \item Distillation on reconstructed feature maps (FitNet) - middle layer strategy \cite{romero2014fitnets}
    \item SimKD with re-used teacher head \cite{SimKD} %Patch to patch distillation framework 
    \item SimKD with [CLS] MLP projector variation
    %\item Structural knowledge distillation
\end{enumerate}
\end{itemize}

\subsection{Datasets}

\begin{table}[ht]
\centering
\caption{Dataset usage for DIC and DLA tasks. 
Symbols: P = pre-training, DP = document pre-training, T = teacher training, S = student training, * = subsampling , E = teacher/student evaluation.}
\begin{tabular}{|l|l|c|c|c|}
\hline
\textbf{Dataset} & \textbf{Task} & \textbf{Usage} & \textbf{Size} & \textbf{\# Cls} \\ \hline
ImageNet \cite{deng2009imagenet} & DIC & P & 1.28M & 1000 \\ \hline
IIT-CDIP \cite{lewis2006building} & DIC & DP, T, S & 11M & / \\ \hline
\tobacco{} \cite{kumar2013unsupervised} & DIC & T, S, E & 3482 & 10 \\ \hline
\rvl{} \cite{harley2015evaluation} & DIC & DP, T, E, * & 400K & 16 \\ \hline
RVL-CDIP-N \cite{larson2022evaluating} & DIC & E & 10K & 16 \\ \hline
\hline
\end{tabular}
\label{tab:DKD-datasets}
\end{table}

For \DC{}, we benchmark results on both \tobacco{} (original train-val-test splits 800-200-2482) and \rvl.  
We also created a subsampled student training set, \rvlone, by randomly selecting 1K training images per class from \rvl. The originally large training size hinders experimentation (longer iteration cycles).
By evaluating the full \rvl{} test set, we provide a fair evaluation of the usefulness of KD methods, while avoiding the cumbersomeness of student fine-tuning on a large dataset. Additionally, we evaluate on RVL-CDIP-N, which is a test set created for testing robustness to covariate shifts.

\todo[inline]{
METHODOLOGY PROBLEM: no open-source data to test document classification with high label cardinality.
An internal dataset is available.
}

\subsubsection{Preliminary results varying KD strategies - CE teacher}

\begin{table*}
\centering
\caption{Preliminary results of different KD strategies benchmarked for Vit-B applied on the \rvl{} datasets. }
\npdecimalsign{.}
\nprounddigits{3}
\begin{tabular}{|c|c|c|c|n{1}{3}n{1}{3}n{1}{3}|} %{|c|c|c|c||ccc|}  
\hline Dataset & Teacher backbone & Student backbone & Method & \text{ACC} & \text{ECE} & \text{AURC}\\  \hline %$\triangle \mathrm{Acc}$
\rvl & ViT-B & -- & Baseline & 0.891697292432311 &	0.047907298836262	& 0.017503164554713 \\
& -- & ViT-S & Baseline &0.853371334283357 & 0.057859288371025 & 0.02963421682364 \\ %0.90844771119278 & 0.01186920564432 & 0.012508771541678 \\
 & -- & ViT-T & Baseline &0.822045551138778 & 0.042795402533132 & 0.040342312570416 \\ \hline %0.900922523063077 & 0.016201313786121 & 0.014282554880225 \\
\hline \rvlone  & ViT-B & \textbf{ViT-S} & Vanilla [$\tau=2.5, \alpha=0.5$] &0.85427135678392 & 0.048633268034969 & {\npboldmath}0.028376044015258 \\
\hline \rvlone  & ViT-B &  & NKD [$\tau=1, \gamma=1.5$] &0.840471011775294 & 0.073632049497662 & 0.035729143933197 \\
\hline \rvlone  & ViT-B &  & MSE &0.854996374909373 & 0.050951811632588 & {\npboldmath}0.028086462460537 \\
\hline \rvlone  & ViT-B &  & SimKD [CLS+MLP]  & {\npboldmath}0.85947148678717 &	{\npboldmath}0.028154179258953 &	0.28744054887619
 \\
\hline \rvlone  & ViT-B &  & SimKD [CNN] & 0.846796169904247	&0.061565487248227	& 0.141136871882953
 \\
\hline \rvlone  & ViT-B &  & FitNet [middle] &0.842646066151654 & 0.140569373835247 & 0.047789833130271 \\
\hline \rvlone  & ViT-B (ImageNet) & ViT-T & FitNet [middle] & 
0.825895647391185	&0.056582912881468	& 0.155618612805129
 \\
\hline\hline \rvlone  & ViT-B & \textbf{ViT-T} & Vanilla [$\tau=2.5, \alpha=0.5$] &  0.824745618640466 & 0.057973819137046 & {\npboldmath}0.03840438493781 \\
\hline \rvlone  & ViT-B &  & NKD [$\tau=1, \gamma=1.5$]& 0.815070376759419 & 0.093624998224256 & 0.045976107157128 \\
\hline \rvlone  & ViT-B &  & MSE & 0.82329558238956 & 0.065716418501806 & 0.039916989542324 \\
\hline \rvlone  & ViT-B &  & SimKD [CLS+MLP] & {\npboldmath}0.829745743643591 & 0.163239800045468 & 0.094989035012674 \\
\hline \rvlone  & ViT-B &  & SimKD [CNN] & 0.829495737393435 & 0.149636502763522 & 0.055740155161562 \\
\hline \rvlone  & ViT-B &  & FitNet [middle] & 0.812345308632716 & 0.153136880692877 & 0.050564425673713 \\
\hline \rvlone  & ViT-B (ImageNet) & ViT-T & FitNet [middle] & 0.805370134253356	&{\npboldmath}0.05986580106555&	0.167935470434519
 \\
\hline
\end{tabular}
\end{table*}

### 2023-09-26 paper/aurc.tex
AURC update for meeting 26/09

\begin{center}{\noindent\Large\bf Details on AURC}\end{center}

\section{Standard Definition of Selective Classification} %g is independent of model parameters w

Given a data set $\left\{\left(x_i, y_i\right)\right\}_{i=1}^N$ of size $N$ with $\left(x_i, y_i\right)$ independent samples from $\mathcal{X} \times \mathcal{Y}$, and given a pair of functions $(f, g)$, where $g: \mathcal{X} \rightarrow \mathbb{R}$ is a confidence scoring function (CSF) and $f(\cdot): \mathcal{X} \rightarrow \mathcal{Y}$ is the classifier, the classification output after failure detection is defined as:
$$
    (f, g)(x):=\left\{\begin{array}{ll}
        f\left(x\right),  & \text { if } g(x) \geq \tau \\
        \text { filter, } & \text { otherwise }
    \end{array}\right.
$$
Filtering ("detection") is triggered when $g(x)$ falls below a threshold $\tau \in \mathbb{R}$.
In order to perform \textit{meaningful} failure detection, a CSF $g(x)$ is required to output high confidence scores for correct predictions and low confidence scores for 'wrong' predictions based on the binary failure label
$$
    y_{\mathrm{w}}(x)=\mathbb{I}\left[y \neq \hat{y}\right],
$$
where for convenience, $f_{k}(x)$ denotes the $k$-th element of the output vector, thus $\hat{y}=\operatorname{argmax}_{y' \in \mathcal{Y}} f_{y'}(x)$ and $\mathbb{I}$ is the identity function (1 for true events and 0 for false events).

\begin{definition}[Confidence Scoring Function (CSF)]
    Any function whose continuous output \textit{aims to} separate a model's failures from correct predictions can be interpreted as a confidence scoring function. \cite{jaeger2023a}.
    \begin{equation}
        g(x) \in \mathbb{R} \text { s.t. } \left\{\begin{array}{ll}
            g(x) \geq \tau, & \text { if } y_{\mathrm{w}}(x)=0 \\
            g(x) < \tau,    & \text { if } y_{\mathrm{w}}(x)=1
        \end{array}\right.
    \end{equation}
\end{definition}

The goal of the CSF is to induce a partial order over points in $\mathcal{X}$, with optimally reflecting true loss monotonicity in the sense that for every two labeled instances $\left(x_1, y_1\right) \sim P(X, Y)$, and $\left(x_2, y_2\right) \sim P(X, Y)$,
$$
    g\left(x_1, \hat{y} \mid f\right) \leq g\left(x_2, \hat{y} \mid f\right) \Longleftrightarrow \operatorname{Pr}_P\left[\hat{y}_w\left(x_1\right) \neq y_1\right] \geq \operatorname{Pr}_P\left[\hat{y}_w\left(x_2\right) \neq y_2\right]
$$
%the confidence of 1 is lower than 2 if if the probability of failure of 1 is higher than 2; where failure is indicated by the true probability of the failure label $y_{\mathrm{w}}(x)=\mathbb{I}\left[y \neq \hat{y}\right]$.

\noindent Common CSFs used in practice \footnote{While it is preferable to have the output domain of g in $[0,1]$ for easier finding of selection thresholds, this is not a strict requirement.} are:
\begin{enumerate}
    \item Maximum softmax probability (MSP): $g(x) = \max_{y' \in \mathcal{Y}} f_{y'}(x)$
    \item Maximum logit (MaxLogit): $g(x) = \max_{y' \in \mathcal{Y}} z_{y'}(x)$, with \textit{logits} $\mathbf{z} \in \mathbb{R}^K$
    \item Entropy: $g(x) = -\sum_{y' \in \mathcal{Y}} f_{y'}(x) \log f_{y'}(x)$
    \item Distance-based measures (e.g., \cite{fisch2022calibrated})
          \begin{itemize}
              \item kNN distance. A 1D outlier score derived from the average distance of $x$ to its $k$ nearest neighbors in the training distribution
          \end{itemize}
          % \item Mutual information (MI): $g(x) = \sum_{y' \in \mathcal{Y}} f_{y'}(x) \log \frac{f_{y'}(x)}{p(y')}$, where $p(y')$ is the prior probability of class $y'$
          % \item Predictive entropy (PE): $g(x) = -\sum_{y' \in \mathcal{Y}} f_{y'}(x) \log \sum_{y'' \in \mathcal{Y}} f_{y''}(x)$
          % \item Predictive entropy reduction (PER): $g(x) = \sum_{y' \in \mathcal{Y}} f_{y'}(x) \log \frac{f_{y'}(x)}{\sum_{y'' \in \mathcal{Y}} f_{y''}(x)}$
          % \item Distance to decision boundary (D2B): $g(x) = \frac{1}{\norm{\mathbf{w}}_2} \left(\mathbf{w}^T \mathbf{x} + b\right)$, where $\mathbf{w}$ and $b$ are the weights and bias of the last layer of the NN
          %https://github.com/ajfisch/calibrated-selective-classification/blob/main/src/data/features.py 
\end{enumerate}

While it is preferable to have the output domain of g in $[0,1]$ for easier finding of selection thresholds, it is not a strict requirement.

\section*{Jordy's AURC}

A crucial component in making a multi-class classifier $f$ selective is the selector function $g$, which originally defined as a CSF, requires a threshold $\tau$ to be defined in order to perform filtering. The threshold $\tau$ is usually defined as a fixed value, or as a function of the confidence score distribution of the training set.
With the objective of measuring a classifier's selective performance, we need to establish risk and coverage measures that are dependent on the threshold $\tau$.

\begin{definition}[Coverage]
    The coverage of a selective classifier $f$ with selector $g$ and threshold $\tau$ is defined as the expectation of selecting from a distribution.
    \begin{equation}
        \mathcal{C}_{\tau}(f, g)=\mathbb{E}_{P_X}\left[g(x) \geq \tau\right]
    \end{equation}
    Put plainly: the percentage of samples above the threshold %n_samples_above_threshold / n_samples
    % probability mass of the non-rejected region in \mathcal{X} 
\end{definition}

\begin{definition}[Risk]
    The risk of a selective classifier $f$ with selector $g$ and threshold $\tau$ is defined as the expected loss of the classifier on the filtered distribution
    %$\mathcal{D}_{\text {in }}^{\text {test }, \times}$, i.e. the expected loss of the classifier on the set of misclassified in-distribution test samples.
    \begin{equation}
        \mathcal{R}_{\tau}(f, g)=\frac{\mathbb{E}_{P_{XY}}\left[\ell\left(y, f(x)\right) \mathbb{I}\left[g(x) \geq \tau\right]  \right]}{\mathcal{C}_{\tau}(f, g)} %
    \end{equation}
    Put plainly: the percentage of errors above the threshold %residuals / n_samples
\end{definition}

To estimate the area under the risk-coverage curve \AURC{}, a multi-threshold list $\theta_T$ with length $T$ is obtained as the unique values of the ascending ranking of all CSF values $g(X)$. The $\mathrm{AUC}$ is then approximated by the sum of the areas of the trapezoids defined by the points $(\mathcal{C}_{\tau}(f, g), \mathcal{R}_{\tau}(f, g))$ for all thresholds $\tau \in \theta_T$.

%Starting point is risk=1-accuracy, coverage=100\%; Ending point= coverage=0\%, risk=0\%.

%(measured on a given distribution, e.g., the validation set)

\section{End-to-End example of AURC}

%Despite its relevance for application, the often final step of failure detection, i.e. the definition of a decision threshold on the confidence score, is often neglected in research. In Appendix D we present an approach that does not require the calibration of scores; optimize coverage, relative to maximal risk

%\section{Detail on integration for AURC}

\section*{Why AURC?}

Using AURC as a primary evaluation metric has the following advantages \cite{ding2020revisiting}.
\begin{enumerate}
    \item when the underlying prediction models are the same, AURC is an effective quality metric to indicate the performance of selective prediction
    \item  when the prediction models are different, which happens a lot in the literature due to the emerging trend of uncertainty-aware training (or swapping out loss functions), using AURC instead of AUPR and AUROC prevents unfair and potential misleading comparison.
    \item AURC can be generalized to distinct tasks with task-specific evaluation metrics ($\ell(y,f(x))$) while AUPR and AUROC cannot.
    \item AURC is an alternative optimization objective to directly maximize the performance of selective prediction and helps to avoid weighing multiple objective terms in related work
    \item AURC directly shows the cost-performance trade-off in selective prediction which is not visible in the conventional ROC curve or PR curve.
\end{enumerate}

\newpage

### 2023-09-26 src/metrics.py

@@ -1,0 +1,553 @@
+# base calibration metric
+
+# https://github.com/JonathanWenger/pycalib/blob/master/pycalib/scoring.py
+# https://github.com/google-research/robustness_metrics/blob/master/robustness_metrics/metrics/uncertainty.py
+#  https://github.com/kjdhfg/fd-shifts
+
+from __future__ import annotations
+import scipy
+import sklearn.utils.validation
+
+from dataclasses import dataclass
+from functools import cached_property
+from typing import Any
+import pandas as pd
+import numpy as np
+import numpy.typing as npt
+from sklearn import metrics as skm
+import evaluate as HF_evaluate
+
+ArrayType = npt.NDArray[np.floating]
+
+
+# TODO: validate numerically stable softmax
+## https://github.com/IML-DKFZ/fd-shifts/blob/main/fd_shifts/analysis/confid_scores.py#L20
+
+# ----------------------------------- general metrics with consistent metric(y_true, p_hat) API -----------------------------------
+
+
+def f1_micro(y_true, p_hat, y_hat=None):
+    if y_hat is None:
+        y_hat = np.argmax(p_hat, axis=-1)
+    return skm.f1_score(y_true, y_hat, average="micro")
+
+
+def f1_macro(y_true, p_hat, y_hat=None):
+    if y_hat is None:
+        y_hat = np.argmax(p_hat, axis=-1)
+    return skm.f1_score(y_true, y_hat, average="macro")
+
+
+# Pure numpy and TF implementations of proper losses (as metrics) -----------------------------------
+
+
+def brier_loss(y_true, p_hat):
+    r"""Brier score.
+    If the true label is k, while the predicted vector of probabilities is
+    [y_1, ..., y_n], then the Brier score is equal to
+    \sum_{i != k} y_i^2 + (y_k - 1)^2.
+
+    The smaller the Brier score, the better, hence the naming with "loss".
+    Across all items in a set N predictions, the Brier score measures the
+    mean squared difference between (1) the predicted probability assigned
+    to the possible outcomes for item i, and (2) the actual outcome.
+    Therefore, the lower the Brier score is for a set of predictions, the
+    better the predictions are calibrated. Note that the Brier score always
+    takes on a value between zero and one, since this is the largest
+    possible difference between a predicted probability (which must be
[…]

### 2023-09-26 paper/draft.tex
AURC update for meeting 26/09
making sanity check
Updates from Overleaf

@@
 
-\section{Example \AURC}
+\section*{Updates 27/09}
+
+\begin{itemize} %tailor to audience
+    \item $f(x)_{k+1}$: write out with $g$, as it is restrictive to have this additional output in $[0,1]$, can be in $\mathbb{R}$; can have its own parameters and features
+    \item Sanity check complete
+    \item Differentiability: 1) indicator function, 2) comparison (>)
+\end{itemize}
+
+% \section{Example \AURC}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.5\textwidth]{figures/aurc_example.png}
+%     \caption{Example of \AURC{} calculation for a 3-class problem with a reject option.}
+%     \label{fig:aurc_example}
+% \end{figure}
 
@@
 By evaluating the full \rvl{} test set, we provide a fair evaluation of the usefulness of KD methods, while avoiding the cumbersomeness of student fine-tuning on a large dataset.
-Additionally, we evaluate on RVL-CDIP-N, which is a test set created for testing robustness to covariate shifts.
+Additionally, we evaluate RVL-CDIP-N, which is a test set created for testing robustness to covariate shifts.
 
@@
 \end{table*}
+
+% Considering standard Neural Networks (NNs), the last layer outputs a vector of real-valued \textit{logits} $\mathbf{z} \in \mathbb{R}^K$, which in turn are normalized using a sigmoid/softmax activation function.
+
+% $\displaystyle \sigma(z) = \frac{1}{1 + \exp^{-z}}$
+% $\displaystyle \text{softmax}(\mathbf{z}) = \frac{{\exp(z)}}{{\sum_{k=1}^{K} \exp(z_k)}}$
+
+% For convenience, $f_{k}(x)$ denotes the $k$-th element of the output vector.
+
+% $\hat{y} = \argmax_{y'\in{\mathcal{Y}}}f_{y'}(X)$ is the top-1 class prediction
+
+% $\hat{p}= \max_{y'\in{\mathcal{Y}}}f_{y '}(X)$ is the associated posterior probability
+% Some interesting distributions are defined:
+
+% \begin{itemize}
+%     \item $\mathcal{D}_{\text {in }}$ denotes the distribution over $\mathcal{X}$ of in-distribution (ID) data
+%     \item 	 $\mathcal{D}_{\text {in }}^{\text {test }, \checkmark}$ and $\mathcal{D}_{\text {in }}^{\text {test }, \times}$ represent the distribution of correct and misclassified ID test samples
+% \end{itemize}
+
+% Now to combine these two measures into a single metric, we need to define a trade-off between risk and coverage.
+This is done by defining a \textit{cost} function $c(\cdot)$, which is a function of the risk and coverage.
+The cost function is defined as the weighted sum of the risk and coverage, where the weights are defined by the user.
+The cost function is defined as:
+% \begin{equation}
+%     c(\mathcal{R}_{\tau}(f, g), \mathcal{C}_{\tau}(f, g))=\alpha \mathcal{R}_{\tau}(f, g)+(1-\alpha) \mathcal{C}_{\tau}(f, g)
+% where $\alpha \in[0,1]$ is a user-defined parameter that controls the trade-off between risk and coverage. The cost function is a function of the threshold $\tau$ and the selector $g$.

### 2023-09-26 paper/draft.tex
unit test for AURC def 2

@@
 \begin{itemize} %tailor to audience
-    \item $f(x)_{k+1}$: write out with $g$, as it is restrictive to have this additional output in $[0,1]$, can be in $\mathbb{R}$; can have its own parameters and features
-    \item Sanity check complete
+\item $f(x)_{k+1}$: written as $g$, as it is restrictive to have this additional output in $[0,1]$, can be in $\mathbb{R}$; can have its own parameters (e.g., MLP) with features derived from any intermediate representation of $x$
+\item Sanity check complete: naive and matrix version
     \item Differentiability: 1) indicator function, 2) comparison (>)
 \end{itemize}
-
 % \section{Example \AURC}
@@
 By evaluating the full \rvl{} test set, we provide a fair evaluation of the usefulness of KD methods, while avoiding the cumbersomeness of student fine-tuning on a large dataset.
-Additionally, we evaluate RVL-CDIP-N, which is a test set created for testing robustness to covariate shifts.
+Additionally, we evaluate on RVL-CDIP-N, which is a test set created for testing robustness to covariate shifts.
 

### 2023-09-27 paper/draft.tex
Updates from Overleaf
small changes

@@
     \item $f(x)_{k+1}$: written as $g$, as it is restrictive to have this additional output in $[0,1]$, can be in $\mathbb{R}$; can have its own parameters (e.g., MLP) with features derived from any intermediate representation of $x$
-\item Sanity check complete: naive and matrix version
+    \item Different AURC implementations (Geifman vs. Jaeger)
+    \item Sanity check complete: naive version with approximation
     \item Differentiability: 1) indicator function, 2) comparison (>)
 \end{itemize}
+
+\begin{todolist}
+\item Vectorized implementation of Definition 2
+\end{todolist}
+
 % \section{Example \AURC}

### 2023-10-17 paper/draft.tex
checking for alpha

@@
     \operatorname{AURC}(f) := \mathbb{E}_{x \sim p_x} \left[  \frac{\mathbb{E}_{(\tilde{x},\tilde{y})\sim p_{xy}}[\ell([f(\tilde{x})]_{1:k},\tilde{y}) \mathbb{I}[f(\tilde{x})_{k+1} > f(x)_{k+1}]]}{\mathbb{E}_{\tilde{x}\sim p_{x}}[ \mathbb{I}[f(\tilde{x})_{k+1} > f(x)_{k+1}]]} \right] \\
-    \approx \frac{1}{n} \sum_{h=1}^n \frac{\frac{1}{n-1} \sum_{i\neq h} \ell([f(x_i)]_{1:k},y_i) \mathbb{I}[f(x_i)_{k+1} > f(x_h)_{k+1}]}{\frac{1}{n-1} \sum_{i\neq h} \mathbb{I}[f(x_i)_{k+1} > f(x_h)_{k+1}]}
+    \approx \frac{1}{n} \sum_{h=1}^n \frac{\frac{1}{n-1} \sum_{i\neq h} \ell([f(x_i)]_{1:k},y_i) \mathbb{I}[f(x_i)_{k+1} > f(x_h)_{k+1}]}{\frac{1}{n-1} \sum_{i\neq h} \mathbb{I}[f(x_i)_{k+1} > f(x_h)_{k+1}]} := \widehat{\mathrm{AURC}}(f) \\
+   =\frac{1}{n} \sum_{h=1}^n \frac{ \sum_{i\neq h} \ell([f(x_i)]_{1:k},y_i) \mathbb{I}[f(x_i)_{k+1} > f(x_h)_{k+1}]}{ \sum_{i\neq h} \mathbb{I}[f(x_i)_{k+1} > f(x_h)_{k+1}]} = \frac{1}{n} \sum_{i=1}^n \alpha_i \ell([f(x_i)]_{1:k},y_i)
 \end{align}
+
+What is the formula for $\alpha_i$?  $\mathcal{O}(n)$ algorithm for computing all $\alpha_i$ given an index $\{\pi_i\}_{1\leq i\leq n}$ that sorts $\{f(x_i)_{k+1}\}_{1\leq i \leq n}$?
+
+J: Informally, $\alpha_i$ should be a value between 1 and n-1, denoting the number of instances that have higher estimated confidence than  $f(x_i)_{k+1}$.
+
+%loss per h, multiplied by a value between 1 and N-1; linear in terms of the loss
+%linear coefficients with respect to some alpha; alpha is a scalar
+%calculate alphas faster than $O(N^2)$$?
+%sort $(O(NlogN)$ and thresholding $O(N)$ 
+%sort by k+1
 
@@
 \item Vectorized implementation of Definition 2
+\item Write out alphas like weights (impl) in Definition 3
 \end{todolist}

### 2023-10-18 src/diff_aurc.py
Add naive alphas implementation

@@ -1,0 +1,60 @@
+import numpy as np
+
+
+def entropy(x):
+    exp_x = np.exp(x)
+    A = np.sum(exp_x, axis=-1)  # sum of exp(x_i)
+    B = np.sum(x * exp_x, axis=-1)  # sum of x_i * exp(x_i)
+    return np.log(A) - B / A
+
+
+def generate_random_data():
+    num_samples = 10
+    p_test = np.random.rand(num_samples, 3)
+    y_test = np.random.choice(3, num_samples)
+    return p_test, y_test
+
+
+def AURC_naive(f_X, g, Y):
+    N = len(Y)
+    outer_sum = 0
+    for h in range(N):
+        num = 0  # risk
+        den = 0  # coverage
+        for i in range(N):  # inner sums
+            if i == h:
+                continue
+            loss = int(f_X[i].argmax(-1) != Y[i])
+            indicator = g(f_X[i]) > g(f_X[h])
+            num += loss * indicator  # empirical risk for h vs. i
+            den += indicator  # empirical coverage for h vs. i
+
+        if den == 0:  # no coverage, no risk?
+            # print('Comparison for h={} has 0 denominator'.format(h)) #which means it is max?
+            continue
+        outer_sum += num / den  # empirical risk for h vs. all i
+    return outer_sum / N  # , risks, coverages
+
+
+def AURC_naive_alphas(f_X, g, Y):
+    N = len(Y)
+    indices_sorted = np.argsort(g(f_X))
+    sorted_f_X = f_X[indices_sorted]
+    sorted_y = Y[indices_sorted]
+    losses = sorted_f_X.argmax(-1) != sorted_y
+
+    final_sum = 0
+    for i in range(N - 1):
+        loss_sum = np.sum(losses[i+1:])
+        final_sum += loss_sum / (N - i - 1)
+
+    return final_sum / N
+
+
+if __name__ == '__main__':
+    g = lambda x: -entropy(x)
+    p_simp, y_simp = generate_random_data()
+    naive = AURC_naive(p_simp, g, y_simp)
+    print(naive)
+    naive_alphas = AURC_naive_alphas(p_simp, g, y_simp)
+    print(naive_alphas)

### 2023-10-19 src/diff_aurc.py
O(N) alphas implementation

@@ -30,15 +30,16 @@ def AURC_naive(f_X, g, Y):
             den += indicator  # empirical coverage for h vs. i
 
         if den == 0:  # no coverage, no risk?
-            # print('Comparison for h={} has 0 denominator'.format(h)) #which means it is max?
             continue
         outer_sum += num / den  # empirical risk for h vs. all i
     return outer_sum / N  # , risks, coverages
 
 
-def AURC_naive_alphas(f_X, g, Y):
+def AURC_naive_alphas_ON2(f_X, g, Y):
     N = len(Y)
     indices_sorted = np.argsort(g(f_X))
+    
+    #assume all are sorted
     sorted_f_X = f_X[indices_sorted]
     sorted_y = Y[indices_sorted]
     losses = sorted_f_X.argmax(-1) != sorted_y
@@ -47,7 +48,18 @@ def AURC_naive_alphas(f_X, g, Y):
     for i in range(N - 1):
         loss_sum = np.sum(losses[i+1:])
         final_sum += loss_sum / (N - i - 1)
+    return final_sum / N
 
+def AURC_naive_alphas_ON(f_X, g, Y):
+    N = len(Y)
+    indices_sorted = np.argsort(-g(f_X)) #descending sort
+    sorted_f_X = f_X[indices_sorted]
+    sorted_y = Y[indices_sorted]
+    final_sum = 0
+    partial_loss = int(sorted_f_X[0].argmax(-1) != sorted_y[0]) # init by largest
+    for i in range(1, N):
+        final_sum += partial_loss / i
+        partial_loss += int(sorted_f_X[i].argmax(-1) != sorted_y[i])
     return final_sum / N
 
 
@@ -56,5 +68,7 @@ if __name__ == '__main__':
     p_simp, y_simp = generate_random_data()
     naive = AURC_naive(p_simp, g, y_simp)
     print(naive)
-    naive_alphas = AURC_naive_alphas(p_simp, g, y_simp)
+    naive_alphas = AURC_naive_alphas_ON2(p_simp, g, y_simp)
+    print(naive_alphas)
+    naive_alphas = AURC_naive_alphas_ON(p_simp, g, y_simp)
     print(naive_alphas)

### 2023-10-19 src/AURC_implementations.py
Cleaned implementations for standardization - added differentiable loss function to be tested

@@ -1,0 +1,160 @@
+import numpy as np
+from argparse import Namespace
+import math
+import time
+import pandas as pd
+
+
+def runtime_wrapper(input_function, *args, **kwargs):
+    start_value = time.perf_counter()
+    return_value = input_function(*args, **kwargs)
+    end_value = time.perf_counter()
+    runtime_value = end_value - start_value
+    # print(f"Finished executing {input_function.__name__} in {runtime_value} seconds")
+    return return_value, runtime_value
+
+
+def entropy(x):
+    exp_x = np.exp(x)
+    A = np.sum(exp_x, axis=-1)  # sum of exp(x_i)
+    B = np.sum(x * exp_x, axis=-1)  # sum of x_i * exp(x_i)
+    return np.log(A) - B / A
+
+
+def generate_random_data():
+    num_samples = 10
+    p_test = np.random.rand(num_samples, 3)
+    y_test = np.random.choice(3, num_samples)
+    return p_test, y_test
+
+
+##### metrics #####
+
+
+def geifman_AURC(f_X, g, Y, plot=False):
+    residuals = f_X.argmax(-1) != Y
+    confidence = g(f_X)
+    curve = []
+    m = len(residuals)
+    idx_sorted = np.argsort(confidence)
+    temp1 = residuals[idx_sorted]
+    cov = len(temp1)
+    acc = sum(temp1)
+    curve.append((cov / m, acc / len(temp1)))
+    for i in range(0, len(idx_sorted) - 1):
+        cov = cov - 1
+        acc = acc - residuals[idx_sorted[i]]
+        curve.append((cov / m, acc / (m - i)))
+    AUC = sum([a[1] for a in curve]) / len(curve)
+    return AUC
+
+
+def official_AURC(f_X, g, Y):
+    "The RC curve is obtained by computing the risk of the coverage from the beginning of g(x) (most confident) to the end (least confident)."
+    incorrect = f_X.argmax(-1) != Y  # instance-level mask
+    g_X = g(f_X)
+    idx_sorted = np.argsort(g_X)  # in ascending format; construct curve from right to left
+
+    coverages, risks = [], []
+    weights = (
+        []
+    )  # just forms some mask of points that were different/used for integration (x/N) --> with percentage of data captured
+
+    # well-defined starting point: risk = 1-accuracy (loss), coverage=100%; threshold=0
[…]

### 2023-10-19 src/diff_aurc.py

@@ -1,74 +1,0 @@
-import numpy as np
-
-
-def entropy(x):
-    exp_x = np.exp(x)
-    A = np.sum(exp_x, axis=-1)  # sum of exp(x_i)
-    B = np.sum(x * exp_x, axis=-1)  # sum of x_i * exp(x_i)
-    return np.log(A) - B / A
-
-
-def generate_random_data():
-    num_samples = 10
-    p_test = np.random.rand(num_samples, 3)
-    y_test = np.random.choice(3, num_samples)
-    return p_test, y_test
-
-
-def AURC_naive(f_X, g, Y):
-    N = len(Y)
-    outer_sum = 0
-    for h in range(N):
-        num = 0  # risk
-        den = 0  # coverage
-        for i in range(N):  # inner sums
-            if i == h:
-                continue
-            loss = int(f_X[i].argmax(-1) != Y[i])
-            indicator = g(f_X[i]) > g(f_X[h])
-            num += loss * indicator  # empirical risk for h vs. i
-            den += indicator  # empirical coverage for h vs. i
-
-        if den == 0:  # no coverage, no risk?
-            continue
-        outer_sum += num / den  # empirical risk for h vs. all i
-    return outer_sum / N  # , risks, coverages
-
-
-def AURC_naive_alphas_ON2(f_X, g, Y):
-    N = len(Y)
-    indices_sorted = np.argsort(g(f_X))
-    
-    #assume all are sorted
-    sorted_f_X = f_X[indices_sorted]
-    sorted_y = Y[indices_sorted]
-    losses = sorted_f_X.argmax(-1) != sorted_y
-
-    final_sum = 0
-    for i in range(N - 1):
-        loss_sum = np.sum(losses[i+1:])
-        final_sum += loss_sum / (N - i - 1)
-    return final_sum / N
-
-def AURC_naive_alphas_ON(f_X, g, Y):
-    N = len(Y)
-    indices_sorted = np.argsort(-g(f_X)) #descending sort
-    sorted_f_X = f_X[indices_sorted]
-    sorted_y = Y[indices_sorted]
-    final_sum = 0
-    partial_loss = int(sorted_f_X[0].argmax(-1) != sorted_y[0]) # init by largest
-    for i in range(1, N):
-        final_sum += partial_loss / i
-        partial_loss += int(sorted_f_X[i].argmax(-1) != sorted_y[i])
-    return final_sum / N
-
-
[…]

### 2023-10-19 src/test_differentiability_AURC.py

@@ -1,0 +1,190 @@
+from collections import OrderedDict
+import numpy as np
+from scipy.special import softmax
+import pickle
+from sklearn.model_selection import train_test_split
+import torch
+
+from AURC_implementations import IMPLEMENTATIONS, runtime_wrapper
+
+##### data generation for testing #####
+
+
+def generate_random_data():
+    num_samples = 10
+    p_test = np.random.rand(num_samples, 3)
+    y_test = np.random.choice(3, num_samples)
+    return p_test, y_test
+
+
+def generate_random_data_with_confidence(N=10, K=3, concentration=1, onehot=False):
+    # N evaluation instances {(x_i,y_i)}_{i=1}^N
+
+    def random_mc_instance(concentration=1, onehot=False):
+        reference = np.argmax(np.random.dirichlet(([concentration for _ in range(K)])), -1)  # class targets
+        prediction = np.random.dirichlet(([concentration for _ in range(K)]))  # probabilities
+        if onehot:
+            reference = np.eye(K)[np.argmax(reference, -1)]
+        return reference, prediction
+
+    y_test, p_test = list(zip(*[random_mc_instance() for i in range(N)]))
+    y_test = np.array(y_test, dtype=np.int64)
+    p_test = np.array(p_test, dtype=np.float32)
+    return p_test, y_test
+
+
+def load_CIFAR_logits(path="../data/resnet110_c10_logits.p"):
+    """
+    This testing script loads actual probabilisitic predictions from a resnet finetuned on CIFAR
+
+    There are a number of logits-groundtruth pickles available @ https://github.com/markus93/torch.NN_calibration/tree/master/logits
+    [Seems to have moved from Git-LFS to sharepoint]
+    https://tartuulikool-my.sharepoint.com/:f:/g/personal/markus93_ut_ee/EmW0xbhcic5Ou0lRbTrySOUBF2ccSsN7lo6lvSfuG1djew?e=l0TErb
+
+    See https://github.com/markus93/NN_calibration/blob/master/logits/Readme.txt to decode the [model_dataset] filenames
+    """
+
+    def unpickle_probs(file, verbose=0, normalize=False):
+        with open(file, "rb") as f:  # Python 3: open(..., 'rb')
+            y1, y2 = pickle.load(f)  # unpickle the content
+
+        if isinstance(y1, tuple):
+            y_probs_val, y_val = y1
[…]

### 2023-10-19 src/AURC_loss.py
separated loss to be used in other projects

@@ -1,0 +1,77 @@
+from collections import OrderedDict
+import numpy as np
+from scipy.special import softmax
+import torch
+
+##### g: confidence scoring functions #####
+
+
+def entropy(x):
+    exp_x = np.exp(x)
+    A = np.sum(exp_x, axis=-1)  # sum of exp(x_i)
+    B = np.sum(x * exp_x, axis=-1)  # sum of x_i * exp(x_i)
+    return np.log(A) - B / A
+
+
+def entropy_torch(x):
+    exp_x = torch.exp(x)
+    A = torch.sum(exp_x, dim=-1)  # sum of exp(x_i)
+    B = torch.sum(x * exp_x, dim=-1)  # sum of x_i * exp(x_i)
+    return torch.log(A) - B / A
+
+
+def top12_margin(x):
+    """top-1 - top-2"""
+    values, _ = torch.topk(x, k=2, dim=-1)
+    if x.ndim == 1:
+        return values[0] - values[1]
+    return values[:, 0] - values[:, 1]
+
+
+def top12_margin_np(x):
+    values = np.sort(x, axis=-1)
+    if x.ndim == 1:
+        return values[0] - values[1]
+    return values[:, 0] - values[:, 1]
+
+
+CSF_dict = OrderedDict(
+    {
+        "msp": lambda x: torch.max(torch.nn.functional.softmax(x, dim=-1), -1)[0],  # maximum softmax probability
+        "entropy": lambda x: -entropy_torch(x),  # negative entropy
+        "margin": lambda x: top12_margin(x),  # top-1 - top-2
+        "msp_np": lambda x: np.max(softmax(x, axis=-1), -1),  # maximum softmax probability
+        "entropy_np": lambda x: -entropy(x),  # negative entropy
+        "margin_np": lambda x: top12_margin_np(x),  # top-1 - top-2
+    }
+)
+
+
+##### differentiable loss #####
+
+
+class AURCLoss(torch.nn.Module):
+
+    """PyTorch version of differentiable AURC (https://arxiv.org/pdf/1805.08206.pdf)
+
+    \operatorname{AURC}(f) := \frac{1}{n} \sum_{i=1}^{n-1} \frac{\sum_{j=i+1}^n\ell([f(x_{i:n})]_{1:k},y_j)}{(n-i)}
+
+    This version has time complexity O(N), yet does not take into account duplicate values in g(f_X)
+    """
+
[…]

### 2023-10-19 src/test_differentiability_AURC.py

@@ -5,10 +5,10 @@ import pickle
 from sklearn.model_selection import train_test_split
 import torch
 
+from AURC_loss import AURCLoss, CSF_dict
 from AURC_implementations import IMPLEMENTATIONS, runtime_wrapper
 
 ##### data generation for testing #####
-
 
 def generate_random_data():
     num_samples = 10
@@ -72,79 +72,6 @@ def load_CIFAR_logits(path="../data/resnet110_c10_logits.p"):
     return (p_val, y_val), (p_test, y_test)
 
 
-##### g: confidence scoring functions #####
-
-
-def entropy(x):
-    exp_x = np.exp(x)
-    A = np.sum(exp_x, axis=-1)  # sum of exp(x_i)
-    B = np.sum(x * exp_x, axis=-1)  # sum of x_i * exp(x_i)
-    return np.log(A) - B / A
-
-
-def entropy_torch(x):
-    exp_x = torch.exp(x)
-    A = torch.sum(exp_x, dim=-1)  # sum of exp(x_i)
-    B = torch.sum(x * exp_x, dim=-1)  # sum of x_i * exp(x_i)
-    return torch.log(A) - B / A
-
-
-def top12_margin(x):
-    """top-1 - top-2"""
-    values, _ = torch.topk(x, k=2, dim=-1)
-    if x.ndim == 1:
-        return values[0] - values[1]
-    return values[:, 0] - values[:, 1]
-
-
-def top12_margin_np(x):
-    values = np.sort(x, axis=-1)
-    if x.ndim == 1:
-        return values[0] - values[1]
-    return values[:, 0] - values[:, 1]
-
-
-CSF_dict = OrderedDict(
-    {
-        "msp": lambda x: torch.max(torch.nn.functional.softmax(x, dim=-1), -1)[0],  # maximum softmax probability
-        "entropy": lambda x: -entropy_torch(x),  # negative entropy
-        "margin": lambda x: top12_margin(x),  # top-1 - top-2
-        "msp_np": lambda x: np.max(softmax(x, axis=-1), -1),  # maximum softmax probability
-        "entropy_np": lambda x: -entropy(x),  # negative entropy
-        "margin_np": lambda x: top12_margin_np(x),  # top-1 - top-2
-    }
-)
-
-
-##### differentiable loss #####
-
-
-class AURCLoss(torch.nn.Module):
-
-    """PyTorch version of differentiable AURC (https://arxiv.org/pdf/1805.08206.pdf)
-
[…]

### 2023-10-19 paper/draft.tex

@@
 
-What is the formula for $\alpha_i$?  $\mathcal{O}(n)$ algorithm for computing all $\alpha_i$ given an index $\{\pi_i\}_{1\leq i\leq n}$ that sorts $\{f(x_i)_{k+1}\}_{1\leq i \leq n}$?
+What is the formula for $\alpha_i$? \\
+$\mathcal{O}(n)$ algorithm for computing all $\alpha_i$ given an index $\{\pi_i\}_{1\leq i\leq n}$ that sorts $\{f(x_i)_{k+1}\}_{1\leq i \leq n}$?
 
-J: Informally, $\alpha_i$ should be a value between 1 and n-1, denoting the number of instances that have higher estimated confidence than  $f(x_i)_{k+1}$.
+\begin{align} % number of samples in between each unique integration point
+    \mathcal{C}_j = %
+    \frac{1}{n-1} 
+    \sum_{\substack{h=1 \\ j \neq h}}^n \mathbb{I}[f(x_h)_{k+1} > f(x_j)_{k+1}] \\ %higher CSF than point j
+    \alpha_i =  \mathcal{C}_{\pi_i} - \mathcal{C}_{\pi_{i+1}}     
+\end{align}
+
+\begin{align}
+    \frac{1}{n} \sum_{i=1}^{n-1} \frac{\sum_{j=i+1}^n\ell([f(x_{i:n})]_{1:k},y_j)}{(n-i)}
+\end{align}
+
+\begin{exmp}[simple]\hspace{0.1cm}
+
+    $f(X)_{k+1}$ = [5, 2, 4]
+
+    $\ell([f(X)]_{1:k}, Y)$ = [0, 1, 0]
+
+    $\pi$ = [1,2,0]
+
+\end{exmp}
+
+\section*{Updates 18/10}
+
+J: as it is implemented in \cite{jaeger2023a}, the alphas/weights denote the number of instances between each contiguous pair of thresholds (in our case based on $f(X)_{k+1}$). 
+If we consider each individual confidence estimate as an integration point, the stepsize between them (i.e. alpha) will be a constant $1/N$.
+
+%WRONG: J: Informally, $\alpha_i$ should be a value between 1 and n-1, denoting the number of instances that have higher estimated confidence than  $f(x_i)_{k+1}$. 
+%The most efficient way to obtain is by sorting $f(X)_{k+1}$ in ascending order ( O(nlogn) average case with comparison-based sort such as timsort [default in Python]), then iterate. 
+%https://stackoverflow.com/questions/13603042/what-is-the-fastest-way-to-get-a-sorted-unique-list-in-python
 
@@
 \end{todolist}
+
+\section*{Updates }
+
+\subsection{Literature review on differentiable \AURC}
+
+\cite{cao2022generalizing} close in spirit, dependent on cost, similarly defines it as a $K+1$ classification problem.
 

### 2023-10-19 src/test_differentiability_AURC.py
Added batchsize consistency vs. population loss

@@ -10,6 +10,7 @@ from AURC_implementations import IMPLEMENTATIONS, runtime_wrapper
 
 ##### data generation for testing #####
 
+
 def generate_random_data():
     num_samples = 10
     p_test = np.random.rand(num_samples, 3)
@@ -72,6 +73,44 @@ def load_CIFAR_logits(path="../data/resnet110_c10_logits.p"):
     return (p_val, y_val), (p_test, y_test)
 
 
+def test_all_options(p_test, y_test):
+    # all options: g, loss function, metric implementations
+    loss_function = torch.nn.CrossEntropyLoss()
+    for CSF, g in CSF_dict.items():
+        if "_np" in CSF:
+            continue
+        loss_fx = AURCLoss(g=g, loss_function=loss_function)
+        loss = loss_fx(p_test, y_test)
+        loss.backward()
+        # print(p_test.grad)
+
+        # test metrics vs. loss function
+        for _, impl in IMPLEMENTATIONS:
+            if args.N > 1000 and impl.__name__ == "naive":
+                continue
+            g = CSF_dict[CSF + "_np"]
+            v, t = runtime_wrapper(impl, p_test.detach().numpy(), g, y_test.numpy())
+            print(f"{CSF}: loss: {loss.item()} vs. {impl.__name__}: {v} taking {round(t,4)}s")
+
+
+def test_consistency(p_test, y_test):
+    def batch(iterable, n=1):
+        length = len(iterable)
+        for ndx in range(0, length, n):
+            yield iterable[ndx : min(ndx + n, length)]
+
+    # random batching with different N
+    for Ns in [2, 10, 20, 50, 100, 1000, 2500]:
+        p_test_bs, y_test_bs = batch(p_test, Ns), batch(y_test, Ns)
+        batched_loss = sum(
+            [AURCLoss()(p_test_batch, y_test_batch) for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
+        )
+        print(f"Batched: N={Ns}, loss: {batched_loss.item()}")
+    # full
+    loss_fx = AURCLoss()
+    loss = loss_fx(p_test, y_test)
+    print(f"Population: loss: {loss.item()}")
+
 
 if __name__ == "__main__":
     # add argparse for N and K
@@ -94,26 +133,12 @@ if __name__ == "__main__":
     else:
         p_test, y_test = generate_random_data_with_confidence(N=args.N, K=args.K, concentration=1, onehot=False)
 
[…]

### 2023-10-19 src/test_differentiability_AURC.py
Batch consistency test done

@@ -9,7 +9,7 @@ from AURC_loss import AURCLoss, CSF_dict
 from AURC_implementations import IMPLEMENTATIONS, runtime_wrapper
 
 ##### data generation for testing #####
-
+np.random.seed(42)
 
 def generate_random_data():
     num_samples = 10
@@ -85,8 +85,8 @@ def test_all_options(p_test, y_test):
         # print(p_test.grad)
 
         # test metrics vs. loss function
-        for _, impl in IMPLEMENTATIONS:
-            if args.N > 1000 and impl.__name__ == "naive":
+        for key, impl in IMPLEMENTATIONS:
+            if args.N > 1000 and key == "naive":
                 continue
             g = CSF_dict[CSF + "_np"]
             v, t = runtime_wrapper(impl, p_test.detach().numpy(), g, y_test.numpy())
@@ -100,10 +100,10 @@ def test_consistency(p_test, y_test):
             yield iterable[ndx : min(ndx + n, length)]
 
     # random batching with different N
-    for Ns in [2, 10, 20, 50, 100, 1000, 2500]:
+    for Ns in [2, 10, 20, 50, 100, 1000, 2500, 5000]:
         p_test_bs, y_test_bs = batch(p_test, Ns), batch(y_test, Ns)
-        batched_loss = sum(
-            [AURCLoss()(p_test_batch, y_test_batch) for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
+        batched_loss = np.mean(
+            [AURCLoss()(p_test_batch, y_test_batch).detach().numpy() for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
         )
         print(f"Batched: N={Ns}, loss: {batched_loss.item()}")
     # full

### 2023-10-24 src/test_differentiability_AURC.py
Batch approximation plots

@@ -9,7 +9,7 @@ from AURC_loss import AURCLoss, CSF_dict
 from AURC_implementations import IMPLEMENTATIONS, runtime_wrapper
 
 ##### data generation for testing #####
-np.random.seed(42)
+
 
 def generate_random_data():
     num_samples = 10
@@ -99,18 +99,51 @@ def test_consistency(p_test, y_test):
         for ndx in range(0, length, n):
             yield iterable[ndx : min(ndx + n, length)]
 
+    #np.random.seed(args.seed)
+    
     # random batching with different N
-    for Ns in [2, 10, 20, 50, 100, 1000, 2500, 5000]:
+    collection = OrderedDict()
+    std_errors = OrderedDict()
+    for Ns in [2, 5, 10, 20, 50, 100, 1000, 2500, 5000, 10000]:
         p_test_bs, y_test_bs = batch(p_test, Ns), batch(y_test, Ns)
-        batched_loss = np.mean(
-            [AURCLoss()(p_test_batch, y_test_batch).detach().numpy() for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
-        )
-        print(f"Batched: N={Ns}, loss: {batched_loss.item()}")
+        if Ns == 3:
+            print(len(list(y_test_bs)[-1]))
+                    
+        batched_losses = [AURCLoss()(p_test_batch, y_test_batch).detach().numpy() for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
+        batched_loss = np.mean(batched_losses)
+        collection[Ns] = batched_loss
+        std_errors[Ns] = np.std(batched_losses) / np.sqrt(len(batched_losses))
+        print(f"Batched: N={Ns}, loss: {batched_loss.item()}, std: {np.std(batched_losses)}")
+        
     # full
     loss_fx = AURCLoss()
     loss = loss_fx(p_test, y_test)
+    collection[20000] = loss.item()
+    std_errors[20000] = 0
     print(f"Population: loss: {loss.item()}")
+    
+    #plot collection over Ns (with std errors) in plotly
+    
+    import plotly.graph_objects as go
+    fig = go.Figure()
+    fig.add_trace(go.Scatter(x=list(collection.keys()), y=list(collection.values()), name="Batched",  error_y=dict(type='data', array=list(std_errors.values()), visible=True)))
[…]

### 2023-10-24 src/test_differentiability_AURC.py
Batch approximation plots v2: multiple seeds

@@ -101,19 +101,28 @@ def test_consistency(p_test, y_test):
 
     #np.random.seed(args.seed)
     
-    # random batching with different N
+    # random batching with different N; now extend to different seeds
     collection = OrderedDict()
     std_errors = OrderedDict()
-    for Ns in [2, 5, 10, 20, 50, 100, 1000, 2500, 5000, 10000]:
-        p_test_bs, y_test_bs = batch(p_test, Ns), batch(y_test, Ns)
-        if Ns == 3:
-            print(len(list(y_test_bs)[-1]))
-                    
-        batched_losses = [AURCLoss()(p_test_batch, y_test_batch).detach().numpy() for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
-        batched_loss = np.mean(batched_losses)
-        collection[Ns] = batched_loss
-        std_errors[Ns] = np.std(batched_losses) / np.sqrt(len(batched_losses))
-        print(f"Batched: N={Ns}, loss: {batched_loss.item()}, std: {np.std(batched_losses)}")
+    NN = [2, 5, 10, 20, 50, 100, 1000, 2500, 5000, 10000]
+    for Ns in NN:
+        collection[Ns] = []
+        std_errors[Ns] = []
+        for seed in NN + list(range(20000, 20200)):
+            permutation = np.random.RandomState(seed=seed).permutation(len(y_test))
+            p_test_bs, y_test_bs = p_test[permutation], y_test[permutation]
+            p_test_bs, y_test_bs = batch(p_test, Ns), batch(y_test, Ns)
+
+            batched_losses = [AURCLoss()(p_test_batch, y_test_batch).detach().numpy() for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
+            batched_loss = np.mean(batched_losses)
+            collection[Ns].append(batched_loss)
+            std_errors[Ns].append(np.std(batched_losses) / np.sqrt(len(batched_losses)))
+            
+        collection[Ns] = np.mean(collection[Ns])
+        std_errors[Ns] = np.mean(std_errors[Ns]) #average std error over seeds
+        
+        print(f"Batched: N={Ns}, loss: {collection[Ns]}, std: {std_errors[Ns]}")
+        
         
     # full
     loss_fx = AURCLoss()

### 2023-10-24 src/test_differentiability_AURC.py
Synthetic experiment with batchsize vs. population

@@ -108,7 +108,7 @@ def test_consistency(p_test, y_test):
     for Ns in NN:
         collection[Ns] = []
         std_errors[Ns] = []
-        for seed in NN + list(range(20000, 20200)):
+        for seed in NN + list(range(20000, 2020)):
             permutation = np.random.RandomState(seed=seed).permutation(len(y_test))
             p_test_bs, y_test_bs = p_test[permutation], y_test[permutation]
             p_test_bs, y_test_bs = batch(p_test, Ns), batch(y_test, Ns)
@@ -118,11 +118,10 @@ def test_consistency(p_test, y_test):
             collection[Ns].append(batched_loss)
             std_errors[Ns].append(np.std(batched_losses) / np.sqrt(len(batched_losses)))
             
-        collection[Ns] = np.mean(collection[Ns])
         std_errors[Ns] = np.mean(std_errors[Ns]) #average std error over seeds
+        collection[Ns] = np.mean(collection[Ns])
         
         print(f"Batched: N={Ns}, loss: {collection[Ns]}, std: {std_errors[Ns]}")
-        
         
     # full
     loss_fx = AURCLoss()
@@ -132,7 +131,6 @@ def test_consistency(p_test, y_test):
     print(f"Population: loss: {loss.item()}")
     
     #plot collection over Ns (with std errors) in plotly
-    
     import plotly.graph_objects as go
     fig = go.Figure()
     fig.add_trace(go.Scatter(x=list(collection.keys()), y=list(collection.values()), name="Batched",  error_y=dict(type='data', array=list(std_errors.values()), visible=True)))

### 2023-10-25 src/AURC_loss.py
zero one loss ablation

@@ -49,6 +49,9 @@ CSF_dict = OrderedDict(
 
 ##### differentiable loss #####
 
+def zero_one_loss(input, target):
+    """0-1 loss (inverse)"""
+    return (torch.argmax(input, dim=-1) != target).float().mean()
 
 class AURCLoss(torch.nn.Module):
 
@@ -70,7 +73,7 @@ class AURCLoss(torch.nn.Module):
         sorted_f_X = input[indices_sorted]
         sorted_y = target[indices_sorted]
         final_sum = 0
-        partial_loss = self.loss_function(sorted_f_X[0], sorted_y[0])
+        partial_loss = self.loss_function(sorted_f_X[0], sorted_y[0]) #initialize by largest (often 0 anyway)
         for i in range(1, B):
             final_sum += partial_loss / i
             partial_loss += self.loss_function(sorted_f_X[i], sorted_y[i])

### 2023-10-25 src/test_differentiability_AURC.py

@@ -5,7 +5,7 @@ import pickle
 from sklearn.model_selection import train_test_split
 import torch
 
-from AURC_loss import AURCLoss, CSF_dict
+from AURC_loss import AURCLoss, zero_one_loss, CSF_dict
 from AURC_implementations import IMPLEMENTATIONS, runtime_wrapper
 
 ##### data generation for testing #####
@@ -93,7 +93,7 @@ def test_all_options(p_test, y_test):
             print(f"{CSF}: loss: {loss.item()} vs. {impl.__name__}: {v} taking {round(t,4)}s")
 
 
-def test_consistency(p_test, y_test):
+def test_consistency(p_test, y_test, loss_function=torch.nn.CrossEntropyLoss()):
     def batch(iterable, n=1):
         length = len(iterable)
         for ndx in range(0, length, n):
@@ -101,6 +101,7 @@ def test_consistency(p_test, y_test):
 
     #np.random.seed(args.seed)
     
+    print(f"Testing consistency of AURC loss with risk fx {loss_function} with different batch sizes")
     # random batching with different N; now extend to different seeds
     collection = OrderedDict()
     std_errors = OrderedDict()
@@ -113,7 +114,7 @@ def test_consistency(p_test, y_test):
             p_test_bs, y_test_bs = p_test[permutation], y_test[permutation]
             p_test_bs, y_test_bs = batch(p_test, Ns), batch(y_test, Ns)
 
-            batched_losses = [AURCLoss()(p_test_batch, y_test_batch).detach().numpy() for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
+            batched_losses = [AURCLoss(loss_function=loss_function)(p_test_batch, y_test_batch).detach().numpy() for p_test_batch, y_test_batch in zip(p_test_bs, y_test_bs)]
             batched_loss = np.mean(batched_losses)
             collection[Ns].append(batched_loss)
             std_errors[Ns].append(np.std(batched_losses) / np.sqrt(len(batched_losses)))
@@ -183,6 +184,7 @@ if __name__ == "__main__":
     p_test = torch.from_numpy(p_test).requires_grad_()
     y_test = torch.from_numpy(y_test)
     
+    test_consistency(p_test, y_test, loss_function=zero_one_loss)
     test_consistency(p_test, y_test)
     
     test_all_options(p_test, y_test)

### 2023-10-25 paper/draft.tex

--- 2023-10-17 paper/draft.tex
@@
 
-What is the formula for $\alpha_i$?  $\mathcal{O}(n)$ algorithm for computing all $\alpha_i$ given an index $\{\pi_i\}_{1\leq i\leq n}$ that sorts $\{f(x_i)_{k+1}\}_{1\leq i \leq n}$?
+What is the formula for $\alpha_i$? <TBD> \\
 
-J: Informally, $\alpha_i$ should be a value between 1 and n-1, denoting the number of instances that have higher estimated confidence than  $f(x_i)_{k+1}$.
+%informal answer: the sum over all losses for samples that are ranked higher than (i), normalized by the rank of i in \pi; but how does its own loss come in?
+NOT --> maybe it is worthwhile to consider AURC + CE
+
+%O(N^2) version with ascending sort
+\begin{align}
+    \frac{1}{n} \sum_{i=1}^{n-1} \frac{\sum_{j=i+1}^n\ell([f(x_{i:n})]_{1:k},y_j)}{(n-i)}\label{eq:aurcloss_n2}
+\end{align}
+
+$\mathcal{O}(n)$ algorithm for computing all $\alpha_i$ given an index $\{\pi_i\}_{1\leq i\leq n}$ that sorts $\{f(x_i)_{k+1}\}_{1\leq i \leq n}$ in descending order.
+
+%O(N) version with descending sort and assumes zero-based numbering
+\begin{align}
+    \frac{1}{n} \sum_{i=1}^{n} \frac{1}{i} \sum_{j=0}^{i-1}\ell([f(x_{\pi_j})]_{1:k},y_{\pi_j})
+    \label{eq:aurcloss_n}
+\end{align}
+
+%if sorted in descending order, then can keep incrementing a partial term containing all losses until i-1; skip first, except for its loss
+
+%version that is very close to the implementation
+
+\section*{Updates 25/10}
+
+\begin{itemize}
+    \item $\mathcal{O}(n)$ algorithm implemented and converted to differentiable loss function $\ell_{\mathrm{AURC}}$
+    \item batch approximation to test consistency
+    \item training with $\ell_{\mathrm{AURC}}$ and test effect of different batch sizes
+    \item Literature review on differentiable \AURC TBD
+    \begin{enumerate}
+        \item \cite{cao2022generalizing} close in spirit, dependent on cost, similarly defines it as a $K+1$ classification problem. 
+    \end{enumerate}
+\end{itemize}
+
+\section*{Updates 18/10}
+
+From the alphas, which are individual loss weights, we derived \Cref{eq:aurcloss_n2}, which motivates higher weights for samples that are ranked higher than they should be.
+
+% J: as it is implemented in \cite{jaeger2023a}, the alphas/weights denote the number of instances between each contiguous pair of thresholds (in our case based on $f(X)_{k+1}$). 
+% If we consider each individual confidence estimate as an integration point, the stepsize between them (i.e. alpha) will be a constant $1/N$.
+
+%WRONG: J: Informally, $\alpha_i$ should be a value between 1 and n-1, denoting the number of instances that have higher estimated confidence than  $f(x_i)_{k+1}$. 
+%The most efficient way to obtain is by sorting $f(X)_{k+1}$ in ascending order ( O(nlogn) average case with comparison-based sort such as timsort [default in Python]), then iterate. 
+%https://stackoverflow.com/questions/13603042/what-is-the-fastest-way-to-get-a-sorted-unique-list-in-python
 
@@
 % \end{figure}
+
+\begin{exmp}[simple]\hspace{0.1cm}
+
+    $f(X)_{k+1}$ = [5, 2, 4]
+
+    $\ell([f(X)]_{1:k}, Y)$ = [0, 1, 0]
+
+    $\pi$ = [1,2,0]
+
+\end{exmp}
 

### 2023-10-25 src/AURC_implementations.py
Matthews alphas version

@@ -140,12 +140,27 @@ def AURC_naive_alphas_ON(f_X, g, Y):
         partial_loss += int(sorted_f_X[i].argmax(-1) != sorted_y[i])
     return final_sum / N
 
+def AURC_MB_alphas(f_X, g, Y):
+    N = len(Y)
+    indices_sorted = np.argsort(g(f_X))
+    sorted_f_X = f_X[indices_sorted]
+    sorted_y = Y[indices_sorted]
+    losses = sorted_f_X.argmax(-1) != sorted_y
+    alphas = np.zeros(N)
+    for i in range(N):
+        alphas[i] = 0
+        for j in range(0, i):
+            alphas[i] += (1/(N- j -1))
+    print((alphas/N)[-10:] )
+    return np.mean(alphas * losses)
+
 
 IMPLEMENTATIONS = [
     ("geifman", geifman_AURC),
     ("jaeger 2023", official_AURC),
     ("naive", AURC_naive),
     ("naive_alphas", AURC_naive_alphas_ON),
+    ('AURC_MB_alphas', AURC_MB_alphas)
 ]
 
 

### 2023-10-25 src/AURC_loss.py

@@ -66,7 +66,7 @@ class AURCLoss(torch.nn.Module):
         super(AURCLoss, self).__init__()
         self.g = g
         self.loss_function = loss_function
-
+    
     def forward(self, input, target):
         B = len(target)
         indices_sorted = torch.argsort(self.g(input), descending=True)

### 2023-10-25 src/test_differentiability_AURC.py

@@ -73,15 +73,18 @@ def load_CIFAR_logits(path="../data/resnet110_c10_logits.p"):
     return (p_val, y_val), (p_test, y_test)
 
 
-def test_all_options(p_test, y_test):
+def test_all_options(p_test, y_test, loss_function = zero_one_loss):
     # all options: g, loss function, metric implementations
-    loss_function = torch.nn.CrossEntropyLoss()
     for CSF, g in CSF_dict.items():
         if "_np" in CSF:
             continue
         loss_fx = AURCLoss(g=g, loss_function=loss_function)
         loss = loss_fx(p_test, y_test)
-        loss.backward()
+        try:
+            loss.backward()
+        except RuntimeError:
+            print(f"{loss_function} not differentiable")
+            
         # print(p_test.grad)
 
         # test metrics vs. loss function
@@ -184,7 +187,7 @@ if __name__ == "__main__":
     p_test = torch.from_numpy(p_test).requires_grad_()
     y_test = torch.from_numpy(y_test)
     
-    test_consistency(p_test, y_test, loss_function=zero_one_loss)
-    test_consistency(p_test, y_test)
+    #test_consistency(p_test, y_test, loss_function=zero_one_loss)
+    #test_consistency(p_test, y_test)
     
     test_all_options(p_test, y_test)

### 2023-11-28 paper/draft.tex

@@
 
-What is the formula for $\alpha_i$? <TBD> \\
+What is the formula for $\alpha_i$?  \\
+
+Matthew's guess (sorted ascending)
+\begin{align}
+\alpha_i = \frac{1}{n} \sum_{j=1}^{i-1} \frac{1}{n-j}
+\end{align}
+The $n$th harmonic number is given by
+\begin{align}
+    H_n := \sum_{k=1}^n \frac{1}{k} = \psi(n+1) + \gamma
+\end{align}
+where $\gamma \approx 0.577$ is the Euler–Mascheroni constant and $\psi(n) := \frac{\Gamma'(n)}{\Gamma(n)}$ is the digamma function.\footnote{\url{https://mathworld.wolfram.com/HarmonicSeries.html}}
+Written in terms of harmonic numbers
+\begin{align}
+    \alpha_i = \frac{1}{n} \left( H_{n-1} - H_{n-i} \right) = \frac{1}{n} \left( \psi(n) - \psi(n-i+1) \right).
+\end{align}
+Setting $i=\beta n$ for some $0<\beta<1$ and taking the limit
+\iffalse
+\begin{align}
+    \lim_{n\rightarrow \infty} H_{n-1} - H_{n - \beta n} \approx& \lim_{n\rightarrow \infty} \ln (n-1) - \ln (1-\beta)n \\
+    =& \lim_{n\rightarrow \infty} \ln (n-1) - \ln n - \ln (1-\beta) =  -\ln (1-\beta) \implies \\
+    \alpha_i \approx& -\ln \left(1 - \frac{i}{n}\right)
+\end{align}
+\todo[inline]{Double check what we get if we derive the limit using the digamma function  in place of the logarithm.  The below derivation is technically more principled than above.}
+\fi
+\begin{align}
+    \lim_{n\rightarrow \infty} H_{n-1} - H_{n-\beta n} =& \lim_{n\rightarrow \infty} \psi(n) - \psi(n-\beta n+1) = \lim_{n\rightarrow \infty} \psi(n) - \psi((1-\beta) n) - \frac{1}{(1-\beta) n} \\
+    =& \lim_{n\rightarrow \infty} \ln(n) - \frac{1}{2n} - \ln((1-\beta) n) + \frac{1}{2(1-\beta)n} - \frac{1}{(1-\beta)n} \label{eq:digammaAsymptoticLnPlus} \\
+    =& \lim_{n\rightarrow \infty}   - \frac{1}{2(1-\beta)n}  - \frac{1}{2n}  -\ln (1-\beta) \\
+    =& -\ln (1-\beta) \implies \\
+    \alpha_i \approx& -\ln \left(1 - \frac{i}{n}\right)
+\end{align}
+where Equation~\eqref{eq:digammaAsymptoticLnPlus} was obtained using the asymptotic result that $\psi(n) \rightarrow \ln n - \frac{1}{2n}$ as $n \rightarrow \infty$.
+
+This looks closely related to Equation (3) of  \url{https://arxiv.org/pdf/1805.08206.pdf }
 
@@
 
+Teodora's guess:
 %O(N^2) version with ascending sort
 \begin{align}
-    \frac{1}{n} \sum_{i=1}^{n-1} \frac{\sum_{j=i+1}^n\ell([f(x_{i:n})]_{1:k},y_j)}{(n-i)}\label{eq:aurcloss_n2}
+    \operatorname{AURC}(f) = \frac{1}{n} \sum_{i=1}^{n-1} \frac{\sum_{j=i+1}^n\ell([f(x_{j})]_{1:k},y_j)}{(n-i)}\label{eq:aurcloss_n2}
 \end{align}
 
-$\mathcal{O}(n)$ algorithm for computing all $\alpha_i$ given an index $\{\pi_i\}_{1\leq i\leq n}$ that sorts $\{f(x_i)_{k+1}\}_{1\leq i \leq n}$ in descending order.
+% with $\alpha_i = \sum_{j=i+1}^n 1 $
+
+$\mathcal{O}(n)$ algorithm for computing all $\alpha_i$ given an index $\{\pi_i\}_{0\leq i\leq n}$ that sorts $\{f(x_i)_{k+1}\}_{0\leq i \leq n}$ in descending order.
 
@@
 \begin{align}
-    \frac{1}{n} \sum_{i=1}^{n} \frac{1}{i} \sum_{j=0}^{i-1}\ell([f(x_{\pi_j})]_{1:k},y_{\pi_j})
+    \frac{1}{n} \sum_{i=1}^{n} \frac{1}{i} \sum_{j=1}^{i-1}\ell([f(x_{\pi_j})]_{1:k},y_{\pi_j})
     \label{eq:aurcloss_n}
@@
     \item training with $\ell_{\mathrm{AURC}}$ and test effect of different batch sizes
-    \item Literature review on differentiable \AURC TBD
+    \item Literature review on differentiable \AURC{} TBD
     \begin{enumerate}

### 2023-11-28 src/AURC_loss.py
cleaning up documentation

@@ -89,6 +89,7 @@ class AlphaAURCLoss(torch.nn.Module):
     \operatorname{AURC}(f) := \frac{1}{n} \sum_{i=1}^{n-1} \alpha_i \ell([f(x_{i:n})]_{1:k},y_j)}
 
     This version has time complexity O(N), yet does not take into account duplicate values in g(f_X)
+    Additionally, it uses the alpha-based approximation of AURC, which is more stable over batch sizes.
     """
 
     def __init__(self, g=CSF_dict["msp"], loss_function=torch.nn.CrossEntropyLoss()):
@@ -100,5 +101,5 @@ class AlphaAURCLoss(torch.nn.Module):
         N = len(target)
         indices_sorted = torch.argsort(self.g(input), descending=False)
         losses = self.loss_function(input[indices_sorted], target[indices_sorted])
-        alphas = torch.Tensor([-np.log(1 - i / N) for i in range(N)], device=losses.device)
+        alphas = torch.Tensor([-np.log(1 - i / N) for i in range(N)]).to(losses.device)
         return torch.mean(alphas * losses)

### 2023-11-28 src/test_differentiability_AURC.py

@@ -174,7 +174,7 @@ if __name__ == "__main__":
         p_test = np.vstack((p_test, p_val))
         args.K = 10
         if args.N != len(y_test):  # 20000
-            sub_n = np.random.choice(list(range(len(y_test))), args.N)  # replace =True if larger
+            sub_n = np.random.choice(list(range(len(y_test))), args.N) # sample without replacement
             p_test, y_test = p_test[sub_n], y_test[sub_n]
         else:
             args.N = len(y_test)
